Some examples in extended logic counting classes

### 2015-08-04 tex/sharp-sigma-one-plus.tex
1ago

@@

We define the vocabulary $\L = \{ S_1,\dots,S_t, \leq \}$, where $S_1,\dots,S_t$ have arity $b_1,\dots,b_t$. Let
\begin{eqnarray*}
\Truc &=& \{\A \mid \A \text{ is an } \L \text{-structure with a finite domain } A \text{ such that} \\
&& \leq \text{ is interpreted as a total order for } A \}.
\end{eqnarray*}
We also define a set of second order variables ${\cal X} = \{ X_i \mid i\in\N \}$ where $X_i$ has arity $a_i$, and for every $n \in \N$ there are infinite variables in ${\cal X}$ of arity $n$. A quantifier-free $\L$-formula is defined by the following grammar:
\begin{eqnarray*}
\varphi &::=& x = y \ \mid \ S_i(x_1,\dots,x_{b_i}), i \in \{1,\dots,t\} \ \mid \ x \leq y \ \mid \\
&& X_i(x_1,\dots,x_{a_i}), i\in\N \ \mid \\ 
&& (\neg \varphi) \ \mid \ (\varphi \wedge \varphi) \ \mid \ (\varphi \vee \varphi),
\end{eqnarray*}
where $x,y$ and $x_i$ are first order variables for every $i$. We now define an {\sc FO}-extended quantifier-free $\L$-formula as follows:
\begin{eqnarray*}
\varphi &::=& \alpha, \alpha \text{ is an FO-formula over } \L  \ \mid \\
&& X_i(x_1,\dots,x_{a_i}), i\in\N \ \mid \ \\
&& (\neg \varphi) \ \mid \ (\varphi \wedge \varphi) \ \mid \ (\varphi \vee \varphi).
\end{eqnarray*}
Let $\Y = (Y_1,\dots,Y_q)$ be a tuple of second-order variables of arity $c_1,\ldots,c_q$, and let $\y$ be a tuple of first order variables. For every $\L$-formula $\psi(\Y,\y)$, we define the function $f_{\psi(\Y,\y)}:\Truc \to \N$ as follows:
\begin{eqnarray*}
f_{\psi(\Y,\y)}(\A) &=& \mid \{ \langle \P, \e \rangle \mid \A \models \psi(\P,\e) \} \mid,
\end{eqnarray*}
for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\P = (P_1,\ldots,P_q)$ is a tuple of predicates of arity $c_1,\ldots,c_q$, $P_i \subseteq A^{c_i}$ for every $i \in \{1,\ldots,q\}$, and $\e$ is a tuple of elements from $\A$.\\

>>>>

\usepackage[textwidth=2cm,textsize=small]{todonotes}
\newcommand{\cristian}[1]{\todo[color=blue!10]{{\bf Cristian:} #1}}
\newcommand{\icristian}[1]{\todo[inline, color=blue!10]{{\bf Cristian:} #1}}
\newcommand{\marcelo}[1]{\todo[color=red!10]{{\bf Marcelo:} #1}}
\newcommand{\imarcelo}[1]{\todo[inline, color=red!10]{{\bf Marcelo:} #1}}
\newcommand{\martin}[1]{\todo[color=green!10]{{\bf Martin:} #1}}
\newcommand{\imartin}[1]{\todo[inline, color=green!10]{{\bf Martin:} #1}}

@@

We define the vocabulary $\L = \{ S_1,\dots,S_t, \leq \}$, where $S_1,\dots,S_t$ have arity $b_1,\dots,b_t$. Let
\begin{eqnarray*}
\Truc &=& \{\A \mid \A \text{ is an } \L \text{-structure with a finite domain } A \text{ such that} \\
&& \leq \text{ is interpreted as a total order for } A \}.
\end{eqnarray*}
We also define a set of second order variables ${\cal X} = \{ X_i \mid i\in\N \}$ where $X_i$ has arity $a_i$, and for every $n \in \N$ there are infinite variables in ${\cal X}$ of arity $n$. A quantifier-free $\L$-formula is defined by the following grammar:
\begin{eqnarray*}
\varphi &::=& x = y \ \mid \ S_i(x_1,\dots,x_{b_i}), i \in \{1,\dots,t\} \ \mid \ x \leq y \ \mid \\
&& X_i(x_1,\dots,x_{a_i}), i\in\N \ \mid \\ 
&& (\neg \varphi) \ \mid \ (\varphi \wedge \varphi) \ \mid \ (\varphi \vee \varphi),
\end{eqnarray*}
where $x,y$ and $x_i$ are first order variables for every $i$. We now define an {\sc FO}-extended quantifier-free $\L$-formula as follows:
\begin{eqnarray*}
\varphi &::=& \alpha, \alpha \text{ is an FO-formula over } \L  \ \mid \\
&& X_i(x_1,\dots,x_{a_i}), i\in\N \ \mid \ \\
&& (\neg \varphi) \ \mid \ (\varphi \wedge \varphi) \ \mid \ (\varphi \vee \varphi).
\end{eqnarray*}
\icristian{Yo primero definiria la clase de ``$F$-extended quantifier-free $\L$-formula'' para una logica F sobre $\L$ y de ahi hablaria de quantifier free formula y FO-extended. Con esto tambien puedes definir altiro $\#\Sigma_1$, $\#\Sigma_0$ (que no estan definidas) y  $\#\Sigma_1[\mbox{{\sc FO}}]$. Esto tambien nos servira despues para otras logicas que queriamos extender.} 
Let $\Y = (Y_1,\dots,Y_q)$ be a tuple of second-order variables of arity $c_1,\ldots,c_q$, and let $\y$ be a tuple of first order variables. For every $\L$-formula $\psi(\Y,\y)$, we define the function $f_{\psi(\Y,\y)}:\Truc \to \N$ as follows:
\begin{eqnarray*}
f_{\psi(\Y,\y)}(\A) &=& \mid \{ \langle \P, \e \rangle \mid \A \models \psi(\P,\e) \} \mid,
\end{eqnarray*}
for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\P = (P_1,\ldots,P_q)$ is a tuple of predicates of arity $c_1,\ldots,c_q$, $P_i \subseteq A^{c_i}$ for every $i \in \{1,\ldots,q\}$, and $\e$ is a tuple of elements from $\A$.\\
